Spatial Analysis

### 2021-06-01 Scripts/spatial_analisis.Rmd
Got a matching grid!

```{r dat_exploded, eval = T, echo = T}

@@

```{r maps_sf, eval = F, echo = T, results = 'hide'}

@@

# Load it!
eez_sf <- st_read(dsn = path_world,
                        layer =file_path_sans_ext(fnam_world)) %>% 
  rename(eez_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  # st_transform(crs = "+proj=eck4") %>% # for tbular plot
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) %>% #0.1
  filter(eez_name == "USA (East Coast)")

@@

# us_state_w <- rgdal::readOGR(dsn = "~/Downloads/FederalAndStateWaters/FederalAndStateWaters.gdb")
state_sf <-  st_read("~/Downloads/FederalAndStateWaters/FederalAndStateWaters.gdb") %>% 
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  st_transform(crs = 4326) # for matching projections

# plot <- state_sf %>%
  # st_simplify(preserveTopology = TRUE, dTolerance = 10000) %>%
  # ggplot()  +
  # geom_sf(aes(col = Jurisdicti))

# ggsave("plot.png",plot)

@@

### Exploring spatial regions (a.k.a  buffers)

```{r}

# Buffer state waters

state_bf = st_buffer(state_sf, 405000)

@@

# p <- ggplot() +
#   geom_sf(data = eez_sf, aes(), fill = NA) +
#   geom_sf(data = state_bf, aes(color = jurisdicti), fill = NA) +
#   geom_sf(data = state_sf, aes(color = jurisdicti), fill = NA)
#

# ggsave("testplotb.png",p)

@@

#Join eez shapefile wih buffer

state_bf_ezz <- st_join(state_bf,eez_sf,join = st_contains)

pp <- ggplot(state_bf_ezz) +
  geom_sf()

ggsave("testplot.png",p)

>>>>

- Using only "Northeast US Fall and Spring surveys for now

```{r dat_exploded, eval = T, echo = F}

@@

```{r maps_sf, eval = T, echo = T, results = 'hide'}

@@

# Load it!
eez_sf <- st_read(dsn = path_world,
                        layer =file_path_sans_ext(fnam_world)) %>% 
  rename(eez_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) %>% #0.1 for paper
  filter(eez_name == "USA (East Coast)")

@@

# us_state_w <- rgdal::readOGR(dsn = "~/Downloads/FederalAndStateWaters/FederalAndStateWaters.gdb")
state_sf <-  st_read("~/Downloads/FederalAndStateWaters/FederalAndStateWaters.gdb") %>% 
  janitor::clean_names() %>% 
  mutate(jurisdicti = str_to_lower(jurisdicti)) %>% 
  filter(jurisdicti %in% States) %>% 
  rename(state = jurisdicti)
  # st_transform(crs = 4326)  # for matching projections
  # st_simplify(preserveTopology = TRUE, dTolerance = 0.1)

@@

### 1. Expand Spatial regions (a.k.a  buffers)

A buffer of *410000* m (410 KM / 221) overshoots the EEZ a bit, but is eventually cropped

```{r buffer, eval = T, echo = T}

# Buffer state waters
state_bf = st_buffer(state_sf, 410000) %>% 
  st_transform(4326) %>% # to match shape
  st_set_crs(4326)

@@

state_bf_plot <- ggplot() +
  geom_sf(data = eez_sf, aes(), fill = NA) +
  geom_sf(data = state_bf, aes(color = state), fill = NA)
#

# ggsave("../Results/Partial/state_waters_buffer.png",state_bf_plot)

@@

state_bf_plot

### 2. Expand grid within buffer

Here we expand a grid within the buffer so we can estimate the catch within

```{r grid_indexing, eval = T, echo = T, message = F, warning = F}

# Create grid of the region
bbox <- c(st_bbox(state_bf))

# Expand the grid
ne_grid <- expand.grid(
    lon = seq(from = bbox["xmin"], to = bbox["xmax"], by = 0.3),
    lat = seq(from = bbox["ymin"], to = bbox["ymax"], by = 0.3)
) %>% 
  rowid_to_column("index")

# ----------- #
# [TEST] Plot all layers
# ----------- #

# Looks good
state_sf %>%
  st_transform(4326) %>% # to match shape
  st_set_crs(4326) %>%
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) %>%
  ggplot() +
  geom_sf(aes(color = state))+
  # geom_sf(data = eez_sf, aes(),fill =NA) +
  # ggplot()+
  geom_tile(data = ne_grid,
            aes(
              x = lon,
              y = lat
            ),
            alpha = 0.2
  )

#### 2.1 Merge grid and buffers

Here we converted the grid to `sf`, merged with the buffered states and filter out everything outside the states polygon

```{r grid_buf_merge, eval = T, echo = T, message = F, warning = F, fig.width = 10}

# ---------------- #
# Convert grid to sf
# ---------------- #
grid_sf <- st_as_sf(ne_grid,
             coords = c("lon", "lat"),
             crs = 4326) %>% 
  st_join(state_bf) %>% 
  filter(!is.na(state))

# Create data frame for future computations
# Note, will be used in next chunk
grid_bf_dt <- as.data.frame(grid_sf) %>%
    select(index,state)
    # group_by(state) %>% 
    # summarise(length(index))

# ---------------- #
# [TEST] 
# Visualization of grid
# ---------------- #

gridExtra::grid.arrange(
  # Overall (overlapping) position
  ggplot(grid_sf) +
    geom_sf(aes(color = state)) +
    geom_sf(data = eez_sf, aes(),fill =NA) + 
    theme(legend.position = ""),
  # Showing each state separatley
  ggplot() +
    geom_sf(data = grid_sf, aes(color = state),size = 0.1) +
    facet_wrap(~state) +
    theme(legend.position = "top"),
  nrow = 1)

### 3. Crop buffers to EEZ

```{r buff_to_eez, eval = T, echo = T, message = F, warning = F}

# I don't know how to undo st_simplify so need to re-load the shapefile
eez_sf <- st_read(dsn = path_world,
                        layer =file_path_sans_ext(fnam_world)) %>% 
  rename(eez_name = Name) %>% 
  st_transform(crs = 4326) %>% # 4326
  filter(eez_name == "USA (East Coast)")

# Get the overlapping segments
grid_eez_sf <- st_intersection(grid_sf,eez_sf)

# Get final df with index
grid_eez_df <- as.data.frame(grid_eez_sf) %>%
  select(state,index)

# Plot to make sure makes sense (Picasso syle)
grid_eez_sf %>% 
  st_simplify(preserveTopology = TRUE, dTolerance = 10000) %>% #0.1 for paper
  ggplot() +
  geom_sf(aes(color = state))
